### 2024-02-14 outputs/paper.rmd
1

---
title: "Explaining Hate Crime and Immigrant Conflict: A New Perspective Study of Mate Competition"
subtitle: "Reproduction of 'Hate Crimes and Gender Imbalances: Fears over MateCompetition and Violence against Refugees'(Dancygier, Egami, Jamal & Rischke, 2022)"
author: Guoguo Lv, Xiyou Wang, Yetao Guo
thanks: "Code and data are available at: https://github.com/wxywxy666/Hate-Crimes-and-Gender-Imbalances. A replication of various aspects in this paper are available at: https://doi.org/10.1111/ajps.1259"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: ""
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
# Load libraries
library(ggplot2)
library(here)
```

# Introduction

<!-- Background / Broader Context -->

Hate crime and gender imbalance have emerged as a pressing issue in today's society, sparking fears over the disruption of stable relationship formation and social cohesion (FBI 5; Sage books) Explained as the criminal act motivated by bias or prejudice towards particular groups such as race, religion, gender, or identity-based characteristics, hate crimes pose a crucial threat to individuals' and communities' safety. Characterized by disparities in gender imbalance in the portion of men to women within a population has garnered attention for social dynamics and interpersonal relationships due to potential implications (shang2022tackling).

The link between hate crime and gender imbalance, with factors interrelated influencing both phenomena, is multifaceted and complex (chakraborti2015hate). Certain groups, especially women and members of marginalized communities, often experience gender-based inequalities and are affected disproportionately by discrimination and acts of violence. Further, whether in the form of skewed sex ratios or imbalance in social opportunities and economic gender imbalances, it can lead to frustrations and resentment feelings among individuals, potentially increasing animosity toward specific groups and increasing hate-motivated likelihood (leote2023female).

Although increased recognition of hate crime and gender imbalance intertwined nature, comprehensive analysis by assessing their intersection remains limited(james2021federal). The current has focused on the aspects isolated from their phenomena, overlooking the intricate interplay between them. Therefore, there is a significant need for studies that explore the relationship between gender imbalance and hate crimes, thus focusing on the mechanism underlying individual well-being and social cohesion implications.

This study aims to cross the gap by investigating and addressing the following questions:

-   what is the relationship between the proportion of excess males in a municipality and the perceived competition for mates among the population?
-   what are the attitudes towards hate crimes against refugees?
-   what is the relationship between mate competition and support for hate crimes against refugees?

Drawing data from different sources, we seek to elucidate complex phenomena between interconnections and their broader societal implications (FBI 5; Sage books). Through empirical and theoretical analysis rigorously, contribute to a better understanding of these factors during hate-motivated behavior and gender imbalances informing evidence-based intervention and initiatives policies seeking to foster more excellent inclusive gender balance and tolerance in society (chakraborti2015hate).

# Data

## Data Source and Methodology

## Attributes

# Results

```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Excess Malesat the Municipality Level and Individuals’Perceived Mate Competition"}
# Load libraries
library(ggplot2)

# Load and clean data for wave 4
d <- readRDS(here::here("inputs/data/fig1"))
d <- lapply(d, function(x) round(x, 2))

# Classify all p-value
p_values_1st_2nd <- d[[3]][c(1, 4, 7)]
p_values_2nd_3rd <- d[[3]][c(2, 5,8 )]
p_values_total <- d[[3]][c(3, 6, 9)]

# Summary of basic data information
summary_data <- data.frame(
  group = rep(c("All", "Male (18-44)", "Male (30-40)"), each = 3),
  tercile = rep(c("1st tercile", "2nd tercile", "3rd tercile"), 3),
  mean = d[[1]],
  se = d[[2]]
)

# Function to add p-value annotations
add_pvalues <- function(plot, p_values_1st_2nd, p_values_2nd_3rd, p_values_total) {
  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[1]), y = 0.4),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[1]), y = 0.4),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[1]), y = 0.4), 
                           vjust = 0, fontface = "italic", size = 5)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[2]), y = 0.5),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[2]), y = 0.5),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[2]), y = 0.5), 
                           vjust = 0, fontface = "italic", size = 5)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[3]), y = 0.7),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[3]), y = 0.7),
                           hjust = 1.4, vjust = 30)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[3]), y = 0.7), 
                           vjust = 0, fontface = "italic", size = 5)

  return(plot)
}

# Initial plot
p <- ggplot(summary_data, aes(x = tercile, y = mean, group = group)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_point() +
  facet_wrap(~group, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Excess Males",
    y = "Proportion Receiving Male Competition",
    group = "Group Name"
  ) +
  theme(
    strip.text = element_text(size = 20)
  )

# Use function to add p-value annotation
add_pvalues(p, p_values_1st_2nd, p_values_2nd_3rd, p_values_total)
```

The chart effectively shows the correlation between the surplus of males in a population and the resultant perceptions of mate competition, which has been divided into three separate cohorts. The first cohort comprises the general population, offering a broad perspective on the trend, then more closely in two specific age groups of males: 18-44 and 30-40. The data from the chart is organized into terciles, which echoes the changing male surplus levels. Using this approach facilitates a better analysis, as it portrays a consistent pattern where the idea of mate competition escalates in relation to the increase in male surplus. It is noteworthy to note that this pattern is most marked within the 30-40 age demographic. Interpreting this information shows a raised sensitivity to demographic shifts within this group. The chart utilizes a clear visual methodology to show these insights. The dots represent each perception of the tercile. The dots have also been positioned strategically to convey the central tendency of the data. P-values with meticulous annotation surround these dots, providing a statistical testament to the importance of the observed trends. These p-values serve as a rigorous metric, emphasizing the reliability of the depicted correlations.

```{r demgraphs, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Proportion of Respondents Who Agree with\nEach of the Four Hate Crime Statements"}
# Load data
points_list <- readRDS(here::here("inputs/data/points"))
se_p_list <- readRDS(here::here("inputs/data/se"))

# Bar names
bar_name <- rep("", 5)
bar_name_u <- c("Only Means\n(List)", "Only Means\n(Direct)", "Justified", "Message", "Condemn")

# Set plot format
par(mfrow = c(2, 2), mar = c(5, 5, 4, 2))

# Loop to make 4 panels
for (i in 1:4) {
  point <- points_list[[i]]
  se_p <- se_p_list[[i]]

  base <- barplot(point, ylim = c(0, 0.3), names.arg = bar_name, 
                  col = c("cornflowerblue", rep("lightblue1", 4)),
                  ylab = "Proportion of respondents",
                  las = 2,
                  main = paste("Wave", i))

  arrows(base, point - 1.96*se_p, base, point + 1.96*se_p, 
         lwd = 2, angle = 90, length = 0.05, code = 3,
         col = adjustcolor("black", alpha.f = 0.7))

  mtext(bar_name_u, at = base, side = 1, line = 2.5, cex = 1)
}
```

This bar chart contains four panels that present the results of the survey on respondents' level of agreement with four different hate crime statements collected in a total of four waves of the survey study. Under the "Estimate from List Experiment," there is only one category titled "Only Means (List)" with a single dark blue bar. This bar represents the proportion of respondents who agreed with a statement when asked indirectly, as part of a list experiment. The proportion is around 0.15. The "Direct Questions" section includes three light blue bars for the categories. Each category represents a different hate crime statement where respondents were asked directly. "Only Means (Direct)" has a proportion slightly above 0.15, similar to the indirect estimate but with direct questioning. "Justified" shows a slightly lower proportion of agreement, just below the 0.15 mark. "Message" has a proportion just above 0.15, indicating a similar level of agreement to the "Only Means (Direct)" statement. "Condemn" displays the highest agreement, with a proportion close to 0.25, suggesting that more respondents agree with a statement that presumably condemns hate crimes. This visual comparison shows differences in respondent agreement levels based on how the questions were asked, with direct questioning generally eliciting higher agreement levels.

```{r cohorttab, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Estimated Effects of MateCompetition on Support for HateCrime"}
# Calculate points and standard errors again for clarity
point <- readRDS(here::here("inputs/data/points"))[[5]]
se <- readRDS(here::here("inputs/data/se"))[[5]]

# Preparation of figure settings
colors <- c("skyblue", "salmon", "lightgreen", "gold", "violet")
pch_values <- c(15, 16, 17, 18, 19)
names_point <- c("Only Means", "Justified", "Message", "Prevent", "Condemn")

par(mar = c(5, 5, 4, 2))

# Initial plot
plot(seq(1:5), point, pch = pch_values, col = colors, ylim = c(-0.05, 0.35), xlim = c(0.5, 5.5), 
     xlab = "", xaxt = "n", ylab = "Estimated Effects", 
     cex.lab = 1.25, cex.axis = 1.25, cex.main = 1.5, cex = 1.5)

segments(seq(1:5), point - 1.96*se, 
         seq(1:5), point + 1.96*se, col = colors, lwd = 2)

Axis(side = 1, at = seq(1:5), labels = names_point, cex.axis = 1.25)

abline(h = 0, lty = 2)

text(2.5, 0.30, "Significant Effects", cex = 0.8)

# Add four arrows
for (i in 1:4) {
  arrows(2.5, 0.29, i, point[i] + 0.02, length = 0.05)
}
```
This graph depicts the relationship between mate competition and the level of support for hate crime through five distinct categories. Each category is represented by a point on the graph, with a vertical line that indicates the confidence interval or error range around the estimate.
The "Only Means" category has a low estimated effect size around 0.15 with a relatively small confidence interval. The "Justified" category shows a slightly higher effect size than "Only Means", also with a small confidence interval. The "Message" category has an estimated effect similar to "Justified". The "Prevent" category shows the highest estimated effect size than the previous ones, around 0.2. Finally, the "Condemn" category has an effect size close to zero, with a confidence interval that crosses the zero line.
The graph is marked with a horizontal dashed line at the zero mark, which is likely indicating the point of no effect. The Significant Effects are marked on the graph above the first four categories, suggesting that these categories' effect size is statistically significant, whereas the remaining category is not explicitly marked as significant.

# Discussion

\newpage

# References