### 2024-02-15 outputs/paper.Rmd
1

---
title: "Explaining Hate Crime and Immigrant Conflict: A New Perspective Study of Mate Competition"
subtitle: "Reproduction of 'Hate Crimes and Gender Imbalances: Fears over MateCompetition and Violence against Refugees'(Dancygier, Egami, Jamal & Rischke, 2022)"
author: Guoguo Lv, Xiyou Wang, Yetao Guo
thanks: "Code and data are available at: https://github.com/wxywxy666/Hate-Crimes-and-Gender-Imbalances. A replication of various aspects in this paper are available at: https://onlinelibrary.wiley.com/doi/epdf/10.1111/ajps.12595"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The study examines the correlation between gender imbalance, specifically the excess of males in a population, and the consequent competition for mates, as well as the attitudes towards hate crimes against refugees. Utilizing data from various sources, it aims to illuminate the complex interplay between mate competition and the endorsement of hate crimes, addressing gaps in current research. The findings suggest that higher male ratios may contribute to increased mate competition and potentially to greater support for hate-motivated acts against refugees. This research underscores the need for policies that promote gender balance and societal tolerance, contributing to the prevention of hate crimes and enhancing social cohesion."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
# Load libraries
library(ggplot2)
library(here)
```

\newpage

# Introduction

Hate crime and gender imbalance have emerged as a pressing issue in today's society, sparking fears over the disruption of stable relationship formation and social cohesion (FBI 5; Sage books) Explained as the criminal act motivated by bias or prejudice towards particular groups such as race, religion, gender, or identity-based characteristics, hate crimes pose a crucial threat to individuals' and communities' safety. Characterized by disparities in gender imbalance in the portion of men to women within a population has garnered attention for social dynamics and interpersonal relationships due to potential implications [@shang2022tackling].

The link between hate crime and gender imbalance, with factors interrelated influencing both phenomena, is multifaceted and complex [@chakraborti2015hate]. Certain groups, especially women and members of marginalized communities, often experience gender-based inequalities and are affected disproportionately by discrimination and acts of violence. Further, whether in the form of skewed sex ratios or imbalance in social opportunities and economic gender imbalances, it can lead to frustrations and resentment feelings among individuals, potentially increasing animosity toward specific groups and increasing hate-motivated likelihood [@leote2023female].

Although increased recognition of hate crime and gender imbalance intertwined nature, comprehensive analysis by assessing their intersection remains limited [@james2021federal]. The current has focused on the aspects isolated from their phenomena, overlooking the intricate interplay between them. Therefore, there is a significant need for studies that explore the relationship between gender imbalance and hate crimes, thus focusing on the mechanism underlying individual well-being and social cohesion implications.

This study aims to cross the gap by investigating and addressing the following questions:

-   What is the relationship between the proportion of excess males in a municipality and the perceived competition for mates among the population?
-   What are the attitudes towards hate crimes against refugees?
-   What is the relationship between mate competition and support for hate crimes against refugees?

Drawing data from different sources, we seek to elucidate complex phenomena between interconnections and their broader societal implications (FBI 5; Sage books). Through empirical and theoretical analysis rigorously, contribute to a better understanding of these factors during hate-motivated behavior and gender imbalances informing evidence-based intervention and initiatives policies seeking to foster more excellent inclusive gender balance and tolerance in society [@chakraborti2015hate].

While the original paper used Stata [@Stata] for data processing and analysis in its replication package, we use R [@R] for all data wrangling and analysis and R packages list [@list] to clean the original data, ggplot2 [@ggplot2] to produce the figures and readstata13 [@readstata13] to read the dta files.

#Data

## Data Source and Methodology

This original data is collected by Rafaela Dancygier, Naoki Egami, Amaney Jamal, and Ramona Rischke, 2022, published in the American Journal of Political Science. This article explores the causes of hate crimes against refugees, public and government attitudes towards immigration and ground-breaking analysis of mate competition. The study raises two hypotheses: first, whether native men in areas with high competition pressure for male mates will view immigrants as a threat, and second, whether these men will view violence as the only way to get government attention.

## Attributes

“survey.dta” is a dataset used to produce survey analyzes in the paper (Figure 2, Figure 3, Figure 4), which contains 53 variables and collects a total of 4 waves of survey data. Most variables are responses to survey questions. For those, we list “Question” and options for “Answer” ranging from Strongly Disagree to Strongly Support, represented by numbers 1-4. We only use data directly related to hate crimes. For example, in Figure 1, we use the MateComp.cont variable to get the mate competition data (numbers 1-4), and then calculate the p-value to get the results.

# Results

```{r fig1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Excess Males at the Municipality Level and Individuals’ Perceived Mate Competition"}
# Load libraries
library(ggplot2)

# Load and clean data for wave 4
d <- readRDS(here::here("inputs/data/fig1data"))
d <- lapply(d, function(x) round(x, 2))

# Classify all p-value
p_values_1st_2nd <- d[[3]][c(1, 4, 7)]
p_values_2nd_3rd <- d[[3]][c(2, 5,8 )]
p_values_total <- d[[3]][c(3, 6, 9)]

# Summary of basic data information
summary_data <- data.frame(
  group = rep(c("All", "Male (18-44)", "Male (30-40)"), each = 3),
  tercile = rep(c("1st tercile", "2nd tercile", "3rd tercile"), 3),
  mean = d[[1]],
  se = d[[2]]
)

# Function to add p-value annotations
add_pvalues <- function(plot, p_values_1st_2nd, p_values_2nd_3rd, p_values_total) {
  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[1]), y = 0.4),
                           hjust = 1.15, vjust = 25, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[1]), y = 0.4),
                           hjust = 1.15, vjust = 20, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[1]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[1]), y = 0.4), 
                           vjust = 0, fontface = "italic", size = 5)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[2]), y = 0.5),
                           hjust = 1.15, vjust = 25, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[2]), y = 0.5),
                           hjust = 1.15, vjust = 20, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[2]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[2]), y = 0.5), 
                           vjust = 0, fontface = "italic", size = 5)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_1st_2nd[3]), y = 0.7),
                           hjust = 1.15, vjust = 25, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "3rd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_2nd_3rd[3]), y = 0.7),
                           hjust = 1.15, vjust = 20, size = 2.7)

  plot <- plot + geom_text(data = subset(summary_data, tercile == "2nd tercile" & group == unique(summary_data$group)[3]), 
                           aes(label = sprintf("pv = %.2f", p_values_total[3]), y = 0.7), 
                           vjust = 0, fontface = "italic", size = 5)

  return(plot)
}

# Initial plot
p <- ggplot(summary_data, aes(x = tercile, y = mean, group = group)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_point() +
  facet_wrap(~group, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Excess Males",
    y = "Proportion Receiving Male Competition",
    group = "Group Name"
  ) +
  theme(
    strip.text = element_text(size = 20)
  )

# Use function to add p-value annotation
add_pvalues(p, p_values_1st_2nd, p_values_2nd_3rd, p_values_total)
```

The chart effectively shows the correlation between the surplus of males in a population and the resultant perceptions of mate competition, which has been divided into three separate cohorts. The first cohort comprises the general population, offering a broad perspective on the trend, then more closely in two specific age groups of males: 18-44 and 30-40. The data from the chart is organized into terciles, which echoes the changing male surplus levels. Using this approach facilitates a better analysis, as it portrays a consistent pattern where the idea of mate competition escalates in relation to the increase in male surplus. It is noteworthy to note that this pattern is most marked within the 30-40 age demographic. Interpreting this information shows a raised sensitivity to demographic shifts within this group. The chart utilizes a clear visual methodology to show these insights. The dots represent each perception of the tercile. The dots have also been positioned strategically to convey the central tendency of the data. P-values with meticulous annotation surround these dots, providing a statistical testament to the importance of the observed trends. These p-values serve as a rigorous metric, emphasizing the reliability of the depicted correlations.

```{r fig2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Proportion of Respondents Who Agree with\nEach of the Four Hate Crime Statements"}
# Load data
points_list <- readRDS(here::here("inputs/data/points"))
se_p_list <- readRDS(here::here("inputs/data/se"))

# Bar names
bar_name <- rep("", 5)
bar_name_u <- c("Only Means\n(List)", "Only Means\n(Direct)", "Justified", "Message", "Condemn")

# Set plot format
par(mfrow = c(2, 2), mar = c(5, 5, 4, 2))

# Loop to make 4 panels
for (i in 1:4) {
  point <- points_list[[i]]
  se_p <- se_p_list[[i]]

  base <- barplot(point, ylim = c(0, 0.3), names.arg = bar_name, 
                  col = c("cornflowerblue", rep("lightblue1", 4)),
                  ylab = "Proportion of respondents",
                  las = 2,
                  main = paste("Wave", i))

  arrows(base, point - 1.96*se_p, base, point + 1.96*se_p, 
         lwd = 2, angle = 90, length = 0.05, code = 3,
         col = adjustcolor("black", alpha.f = 0.7))

  text(base, par("usr")[3] - 0.03, labels = bar_name_u, srt = 45, adj = 1, xpd = TRUE, cex = 1)
}
```

This bar chart contains four panels that present the results of the survey on respondents' level of agreement with four different hate crime statements collected in a total of four waves of the survey study. Under the "Estimate from List Experiment," there is only one category titled "Only Means (List)" with a single dark blue bar. This bar represents the proportion of respondents who agreed with a statement when asked indirectly, as part of a list experiment. The proportion is around 0.15. The "Direct Questions" section includes three light blue bars for the categories. Each category represents a different hate crime statement where respondents were asked directly. "Only Means (Direct)" has a proportion slightly above 0.15, similar to the indirect estimate but with direct questioning. "Justified" shows a slightly lower proportion of agreement, just below the 0.15 mark. "Message" has a proportion just above 0.15, indicating a similar level of agreement to the "Only Means (Direct)" statement. "Condemn" displays the highest agreement, with a proportion close to 0.25, suggesting that more respondents agree with a statement that presumably condemns hate crimes. This visual comparison shows differences in respondent agreement levels based on how the questions were asked, with direct questioning generally eliciting higher agreement levels.

```{r fig3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Estimated Effects of MateCompetition on Support for HateCrime"}
# Calculate points and standard errors again for clarity
point <- readRDS(here::here("inputs/data/points"))[[5]]
se <- readRDS(here::here("inputs/data/se"))[[5]]

# Preparation of figure settings
colors <- c("skyblue", "salmon", "lightgreen", "gold", "violet")
pch_values <- c(15, 16, 17, 18, 19)
names_point <- c("Only Means", "Justified", "Message", "Prevent", "Condemn")

par(mar = c(5, 5, 4, 2))

# Initial plot
plot(seq(1:5), point, pch = pch_values, col = colors, ylim = c(-0.05, 0.35), xlim = c(0.5, 5.5), 
     xlab = "", xaxt = "n", ylab = "Estimated Effects", 
     cex.lab = 1.25, cex.axis = 1.25, cex.main = 1.5, cex = 1.5)

segments(seq(1:5), point - 1.96*se, 
         seq(1:5), point + 1.96*se, col = colors, lwd = 2)

Axis(side = 1, at = seq(1:5), labels = names_point, cex.axis = 1)

abline(h = 0, lty = 2)

text(2.5, 0.30, "Significant Effects", cex = 0.8)

# Add four arrows
for (i in 1:4) {
  arrows(2.5, 0.29, i, point[i] + 0.02, length = 0.05)
}
```

This graph depicts the relationship between mate competition and the level of support for hate crime through five distinct categories. Each category is represented by a point on the graph, with a vertical line that indicates the confidence interval or error range around the estimate. The "Only Means" category has a low estimated effect size around 0.15 with a relatively small confidence interval. The "Justified" category shows a slightly higher effect size than "Only Means", also with a small confidence interval. The "Message" category has an estimated effect similar to "Justified". The "Prevent" category shows the highest estimated effect size than the previous ones, around 0.2. Finally, the "Condemn" category has an effect size close to zero, with a confidence interval that crosses the zero line. The graph is marked with a horizontal dashed line at the zero mark, which is likely indicating the point of no effect. The Significant Effects are marked on the graph above the first four categories, suggesting that these categories' effect size is statistically significant, whereas the remaining category is not explicitly marked as significant.

# Discussion

##Finding

Figure \@ref(fig:fig1) illustrates a positive correlation between the number of excess males in a population and the perception of mate competition, especially among men aged 30-40. This demographic trend imply that there is an increased sense of rivalry for potential mates in areas with more males. While the graph does not directly discuss hate crimes, the concept ties into broader social issues where such demographic imbalances may heighten competitive behaviors, potentially influencing social tensions and actions, including hate crimes. In Figure \@ref(fig:fig2), two types of responses are depicted for the statement "Only Means (List)"—one from the list experiment and one from direct questioning. The bar representing the list experiment is indeed lower than that for the direct questioning for the statement "Only Means". This suggest that when individuals are provided with a method that offers anonymity and reduces the pressure to conform to social norms, they might be less likely to agree with statements supporting hate crime, contrary to what might be expected. When considering the impact of mate competition on attitudes towards hate crimes Figure \@ref(fig:fig3), it's crucial to consider the evolutionary psychology perspective, which suggests that competition for mates can increase intergroup hostility and aggression [@Buss1997]. These behaviors may be exacerbated in environments with a skewed sex ratio, as individuals may perceive greater competition for potential partners, heightening in-group biases and out-group derogation [@Pedersen1991].

##The Interplay of Demographics and Sociocultural Factors

In order to get a better idea of these relationships, what can be examined is the influences of demographic factors and social-cultural attitudes which trigger hate crimes.

##Masculinity

Masculinity, a cultural story where the dominant and controlling is the man, can be the reason for such attitudes [@Connell2005]. Men are featured in traditional masculinity stories as defenders who are also providers, who are strong physically, violent and who also dominate sexually. These cultural stories greatly influence men's perception of their social status and how they perceive that status in their behaviors and actions towards other people. Often, in communities where masculinity is equated with sexual conquest and access to women [@Baumeister2004], the superfluous males suffer from psychological stress and anxiety, which might be the cause of aggression and hostility towards other groups. This hostility serves as a signal of intergroup conflict, and can be directed at both out-group and in-group members, as part of the competition. Similarly, the societal norms of conforming to these masculine ideals can cause greater anxiety and stress among men who feel they do not meet these standards. In communities where masculinity is strongly tied to the number of women they are able to conquer, men who are short of potential partners may experience increased psychological stress. This stress can be expressed in the form of violence and hostility towards other people. Therefore, an increase in the number of hate crimes is a possible outcome.

##Digital media and educational inequality

Furthermore, it is crucial to consider the following elements: digital media, educational inequality, and masculinity among all ages and social classes. In the age of digital media, people have their minds conformed to behave in unique ways as no other time has witnessed. This is all the more so in the case of social media where assortment of ideas about masculinity and cultural attitudes grows stronger and gets challenged or reinforced there. While ruling on one side, the digital trend can give birth to discussions that are more diversified and add new elements to the already established notions. They will also be recruitment grounds for crowd them to access and spread extreme ideas as well as hate speech, which in return will just speed up the polarization of public opinion and the violence will come to be the niche of the manhood. The digital realm has two different effects on social viewpoints, which could intensify the negative effect of demographic imbalances on the decision-making process of choosing a partner and the subsequent social conflicts.

##Educational inequalities

Educational inequalities, as well, are of great importance when it comes to the development of the public opinion concerning hate crimes. In several societies, the socioeconomic status, race, and place of a person are important factors that decide how much education they get that is of high quality. However, these differences are the factors that make people perceive the world in different ways and the extent to which they can stand up to the negative stereotypes and prejudices. The people who lack a complete education that provides equal opportunities for girls and boys, teach about social justice and the importance of diversity, are more likely to have biased and discriminating views. Additionally, it is possible that violent displays of masculinity and intolerance will continue to exist as a consequence of education that either promotes traditional gender standards or minimizes the relevance of critical media literacy.

##Implications

On the implication of the findings, it is evident that programs to eliminate support for hate crimes should include both individual attitude change and the broader context, given that the focus is on demographics and society. This covers those promotion of public health that is gender-based and specializes in changing the cultural modeling around masculinity [@Courtenay2000], as well as those reconstructors of social policies envisaging the reduction of economic and social inequalities that can amplify the competition of mates [@Wilkinson2009].
\newpage

# References