### 2024-02-15 outputs/paper.Rmd
final

git config --global user.name "wxywxy666"
git config --global user.email "wxy1700774887@outlook.com"
git config --global --list

gitcreds_set()
create_github_token()

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